Wigglyhedra

### 2024-06-27 draft/wigglyhedra.tex
picture

\begin{figure}
\centerline{
	\includegraphics[scale=1.3]{curve1}
	\qquad
	\includegraphics[scale=1.3]{curve2}
	\qquad
	\includegraphics[scale=1.3]{curve3}
}
\caption{An illustration of the Khovanov--Seidel curve model.}
\label{fig:curveModel}
\end{figure}

\vincent{Add a picture summarizing this section... Asilata drafts it, I will make it.}
In this section we briefly recall how to recover an object of \(\mathcal{C}_{n+1}\) from a wiggly arc on \(n+2\) points on a line.
In fact, the procedure works on more general arcs as well --- namely, those that are not necessarily \(x\)-monotone.

@@

@@

However, the grading convention in~\cite{kho.sei:02} is not the same as ours --- our grading is by path length, while the grading in~\cite{kho.sei:02} is one for arrows in one direction and zero for arrows in the opposite direction.
So we must modify their argument slightly.
In particular, the table in~\cite[Lemma 3.20]{kho.sei:02} that computes the bigraded intersection numbers will change, and so will the table in~\cite[Lemma 4.12]{kho.sei:02} that computes the Poincar{\'e} polynomials of \(\Hom(P_k,-)\).
  \asilata{TODO: this is very tedious; how much to write it?}
\end{proof}
\begin{remark}
  \cref{prop:ext1} is very similar to the condition of being compatible in a cluster-tilting object of a cluster category (see, e.g.~\cite[Cor.~4.3]{bua.mar.rei.ea:06}).
  However, it depends on shifting the objects \(X\) and \(X'\) so that they both lie in the standard heart, which is an auxiliary step that is not required in the case of cluster algebras.
\end{remark}
\begin{remark}
  As stated,~\cref{prop:ext1} in combination with~\cref{prop:hn-filtration} is not expected to be true in greater generality.
  For instance, for the zigzag algebra of Dynkin type~\(D_4\), one can find two indecomposable spherical objects with a non-trivial \(\Ext^1\) that nevertheless appear together in the cohomology filtration of another spherical object.
  However, we expect to see similar constraints on compatibility, which we hope to turn to in forthcoming work.
\end{remark}
\begin{question}
  Suppose we have a collection of spherical objects in the standard heart \(\mathcal{A}_m\) that satisfy the compatibility condition of~\cref{prop:ext1}.
  That is, the \(\Ext^1\) between any pair of objects in this collection is zero.
  \begin{enumerate}
  \item Can we prove properties about a maximal such collection without reference to arcs?
  \item In particular, can we understand the combinatorics of maximal such collections without reference to arcs?
  \item Given a maximal compatible collection of objects, can we always find a larger spherical object such that the factors of the cohomology filtration of that object are precisely shifts of the objects in the given collection?
  \end{enumerate}
\end{question}

>>>>

\begin{figure}
\centerline{
	\begin{tabular}{c@{\qquad}c@{\qquad}c}
		Bounded homotopy category 
		&
		&
		\\
		of graded projective modules 
		&
		$\smash{\xrightarrow[\quad\text{combinatorial model}\quad]{}}$
		&
		curves
		\\
		over the zigzag algebra
		&
		&
		\\
		COMPLEX
		&
		&
		\includegraphics[scale=1.3]{curve1}
		\\[.5cm]
		cohomology $\Bigg\downarrow$ filtration
		&
		&
		$\Bigg\downarrow$ cut
		\\[.5cm]
		Heart of the standard $t$-structure
		&
		$\xrightarrow[\quad\text{combinatorial model}\quad]{}$
		&
		wiggly arcs
		\\
		COMPLEX1
		&
		&
		\includegraphics[scale=1.3]{curve2}
		\\
		COMPLEX2
		&
		&
		\includegraphics[scale=1.3]{curve3}		
	\end{tabular}
}
\caption{An illustration of the Khovanov--Seidel curve model and \cref{prop:cohomology-pieces}.}
\label{fig:curveModel}
\end{figure}

\vincent{Add a picture summarizing this section... Asilata drafts it, I will make it.}
In this section we briefly recall how to recover an object of \(\mathcal{C}_{n+1}\) from a wiggly arc on \(n+2\) points on a line.
In fact, the procedure works on more general curves as well --- namely, those that are not necessarily \(x\)-monotone.

@@

\begin{remark}
Note that in \cref{def:wigglyComplex}, we adopted the convention that the faces of the wiggly complex contain none of the two external wiggly arcs.
In contrast, the wiggly dissections of \cref{prop:hn-filtration} contain at most one of the two external wiggly arcs.
Combinatorially, this is just the Cartesian product of the wiggly complex with an edge.
\end{remark}

@@

However, the grading convention in~\cite{kho.sei:02} is not the same as ours --- our grading is by path length, while the grading in~\cite{kho.sei:02} is one for arrows in one direction and zero for arrows in the opposite direction.
So we must modify their argument slightly.
In particular, the table in~\cite[Lemma 3.20]{kho.sei:02} that computes the bigraded intersection numbers will change, and so will the table in~\cite[Lemma 4.12]{kho.sei:02} that computes the Poincar{\'e} polynomials of \(\Hom(P_k,-)\).
  \asilata{TODO: this is very tedious; how much to write it?}
\end{proof}
\begin{remark}
  \cref{prop:ext1} is very similar to the condition of being compatible in a cluster-tilting object of a cluster category (see, \eg~\cite[Cor.~4.3]{bua.mar.rei.ea:06}) or in a $\tau$-tilting object (see, \eg~\cite{AdachiIyamaReiten}).
  The condition in terms of crossings of curves already appears for some specializations (see, \eg~\cite{FominShapiroThurston, PaluPilaudPlamondon-surfaces}).
  However, in \cref{prop:ext1}, it depends on shifting the objects~\(X\) and~\(X'\) so that they both lie in the standard heart, which is an auxiliary step that is not required in cluster-tilting and $\tau$-tilting theory.
\end{remark}
\begin{remark}
  As stated,~\cref{prop:ext1} in combination with~\cref{prop:hn-filtration} is not expected to be true in greater generality.
  For instance, for the zigzag algebra of Dynkin type~\(D_4\), one can find two indecomposable spherical objects with a non-trivial \(\Ext^1\) that nevertheless appear together in the cohomology filtration of another spherical object.
  However, we expect to see similar constraints on compatibility, which we hope to turn to in forthcoming work.
\end{remark}
\begin{question}
  Suppose we have a collection of spherical objects in the standard heart \(\mathcal{A}_m\) that satisfy the compatibility condition of~\cref{prop:ext1}.
  That is, the \(\Ext^1\) between any pair of objects in this collection is zero.
  \begin{enumerate}
  \item Can we prove properties about a maximal such collection without reference to arcs?
  \item In particular, can we understand the combinatorics of maximal such collections without reference to arcs?
  \item Given a maximal compatible collection of objects, can we always find a larger spherical object such that the factors of the cohomology filtration of that object are precisely shifts of the objects in the given collection?
  \end{enumerate}
\end{question}
